Search-based Synthesis of Network Intrusion Detection Rules

### 2019-09-24 paper/paper.tex
pass on workflow

The SYN Flood attack exploits a vulnerability in the TCP/IP handshake
to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
works as follows. First, a client sends a SYN packet to the server,
requesting a connection. Second, the server responds with a SYN-ACK
packet to the client. Third, the client responds with an ACK message
and the connection is established. Aware of the protocol, an attacker
sends multiple SYN packets to different ports of a server, often using
fake IP addresses.\Mar{Are you saying that there is no timer that the
  server starts when it sends a SYN-ACK message?} Without proper protection, the server accepts the
connection requests and eventually legitimate connection requests
cannot be satisfied due to resource exhaustion.

@@

\tname{} is a search-based approach that uses positive and negative
network traffic examples to synthesize signature-based NIDS rules,
such as those from Snort and Suricata. \tname{} synthesizes rules in
two steps. First, it produces a potentially overspecified rule that
captures the negative example that is provided as input. Then, it uses
positive examples to eliminate unnecessary options from the rule. The
goal of \tname\ is to create a rule that captures only the negative traffic.

>>>>

The SYN Flood attack exploits a vulnerability in the TCP/IP handshake
to establish a TCP connection~\cite{cloudfare-synflood}. The handshake
works as follows. First, a client sends a SYN packet to the server,
requesting a connection. Second, the server responds with a SYN-ACK
packet to the client. Third, the client responds with an ACK message
and the connection is established. Aware of the protocol, an attacker
sends multiple SYN packets to different ports of a server, often using
fake IP addresses. Then, after the server responds with a SYN-ACK
packet, the client sends another SYN to avoid the connection to time
out. Without proper protection, the server accepts the connection
requests and eventually legitimate connection requests cannot be
satisfied due to resource exhaustion.

@@

\begin{figure*}[t]
\centering
\includegraphics[trim=60 400 65 0,clip,width=0.7\textwidth]{figs/nids-workflow}
\caption{\tname\ workflow.}
\label{fig:overview}
\end{figure*}

\tname{} is a search-based approach that uses positive and negative
network examples to synthesize signature-based NIDS rules, such as
those from Snort and Suricata.

The technique synthesizes rules in two steps. First, it produces a
potentially overspecified rule that captures the negative example that
is provided as input. Then, it uses positive examples to eliminate
unnecessary options from the rule. The goal of \tname\ is to create a
rule that only captures the negative traffic.

Figure~\ref{fig:overview} shows the workflow of \tname{} as a pipeline
of two components, each implementing one of the steps mentioned above.
The goal of the first component is to produce a rule that captures the
negative traffic.  First, it initializes the rule encoding with random
values. Only options associated with the used protocol are included in
the encoding. Then, \tname\ optimizes the rule options until \suri\ is
able to capture the attack.  Note that, as all options need to be
satisfied for the traffic to be captured, at this point, any subset of
options would capture the negative traffic. The second component of
the pipeline is initialized with the rule produced in the previous
step. That rule captures the negative traffic but could also capture
benign traffic. This component systematically discards options from
the rule encoding until no positive traffic is captured.

\Fix{describes negative and positive traffic}
